\FloatBarrier
\section{Related Work}

\paragraph{Zero-shot object navigation with foundation models.}
Object-goal navigation \citep{batra2020objectnav} and vision-and-language
navigation \citep{anderson2018r2r,krantz2020vlnce,khanna2024goat} are
typically learned with reinforcement or imitation \citep{wijmans2020ddppo,
ramrakhya2023pirlnav}, which generalizes poorly to unseen distributions.
Recent zero-shot systems instead build on foundation models. One family
guides frontier exploration with language priors
\citep{chaplot2020semexp,zhou2023esc,yu2023l3mvn}; another fuses
open-vocabulary detectors into a value map for goal reasoning
\citep{yokoyama2024vlfm,gadre2023cow}; a third lets a VLM read a panoramic
observation and directly emit a heading \citep{zhou2024navgpt,
long2024instructnav}. We build on InstructNav \citep{long2024instructnav}, a
representative agent that couples a local VLM, an open-vocabulary detector
\citep{wu2024glee}, and a value-map frontier explorer. These works assume
clean observations; we instead study, on the same backbone, \emph{where}
such an agent breaks when its input is degraded.

\paragraph{Robustness of embodied agents to visual corruption.}
RobustNav \citep{chattopadhyay2021robustnav} first benchmarked navigation
under visual and dynamics corruptions and reported large drops, motivating
task-oriented data augmentation \citep{maksymets2021thda} and domain
adaptation. A common alternative is to restore the input before it reaches
the policy via low-light enhancement \citep{guo2020zerodce,wu2023uretinex}
or denoising. These approaches operate at the pixel layer and aim to
recover clean appearance. We show that, for a VLM-driven navigator,
restoring appearance does not restore navigation: an $8.4$\,dB PSNR gain
leaves SR within seed noise, and a per-frame oracle that always picks the
better-looking view does no better. Our benchmark is designed to expose
this gap, with a paired protocol and severity sweeps in the style of
corruption robustness benchmarks \citep{hendrycks2019benchmarking}.

\paragraph{Uncertainty, commitment, and executive control.}
Knowing when a model should not be trusted is central to safe autonomy
\citep{abdar2021uncertainty}; blind image-quality assessment estimates
degradation without a reference \citep{su2020blindly} and active perception
decides where to look in order to reduce uncertainty
\citep{ramakrishnan2021exploration}. A complementary direction studies the
\emph{executive} layer of zero-shot ObjectNav: \emph{when} to commit, stop,
and recover from a bad commitment. Our diagnosis is consistent with this
literature but localizes the problem more sharply: we use a training-free
per-step reliability estimate purely as a \emph{measurement instrument},
and we evaluate deployable commitment-timing and temporal/multi-view
aggregation interventions head-on, each of which leaves success unchanged
or lowers it on a frozen, clean-trained backbone (Tab.~\ref{tab:sweep}).

\paragraph{Evaluation of stochastic agents.}
VLM-driven agents are non-deterministic, and single-run point estimates of
embodied performance are known to be unreliable
\citep{anderson2018evaluation}. We therefore report multi-seed pooled
success, paired per-episode differences, McNemar's exact test, and
bootstrap confidence intervals at $N{=}300$, and treat small-sample point
estimates only as pilot runs; the Experiments section quantifies how badly
small samples mislead in this setting. Figure~\ref{fig:teaser}(e)
visualizes the resulting coverage map of our paper against the families
above.
