\section{Why Navigation Fails Under Degradation}
\label{sec:motivation}

A single episode previews the phenomenon this section quantifies. Asked to
find a toilet under low light, the baseline agent navigates to within
$0.02$\,m of it, never declares it found, wanders off, and the episode ends
in failure (Figure~\ref{fig:case-studies}, top~c). The agent saw enough
to get there; what failed was the decision to stop. This section shows the
pattern is systematic. Every quantity is measured on the unmodified
backbone, organized as six questions (M0--M5): whether degradation hurts
at all (M0), whether the cause is appearance (M1), what failures look
like (M2), whether reliability is informative about commitments (M3),
what triggers a commitment (M4), and whether the mechanism survives at
the $N{=}300$ main-split scale (M5).

\paragraph{Setup.}
We run InstructNav on HM3D ObjectNav \texttt{val\_mini} pooled over seeds
$\{0,1,2\}$ ($n{=}90$, bootstrap $95\%$ CIs, $B{=}5000$). The primary
degradation is low-light at severity~$4$, drawn from a $13$-type
ImageNet-C-style suite \citep{hendrycks2019benchmarking}. The $n{=}90$
study below establishes the qualitative chain; the Experiments section
re-measures decisive quantities at $N{=}300$.

\paragraph{M0--M1: Degradation hurts, but appearance is not the lever.}
Low-light lowers SR from $0.433$ to $0.222$ at $n{=}90$; the paired drop is
$-0.211$ with CI $[-0.322, -0.100]$, excluding zero, so we use low-light as
the primary condition. Figure~\ref{fig:motivation}a shows \emph{where} the drop
lives at the paired $N{=}300$ scale: the per-episode minimum distance to
the goal is nearly unchanged (median $0.11\to0.15$\,m)---the agent still
gets to the goal---yet $48$ episodes flip to failure. Adding a test-time
restoration arm raises PSNR on
low-light from $7.4$ to $15.9$\,dB ($+8.4$\,dB) yet SR moves only $0.233
\to 0.300$ (seed-$0$ slice), inside the $\pm0.10$ seed-noise band:
making the image look clean does not make the agent navigate.

\paragraph{M2: Failures are irreversible commitments.}
To characterize the failures, we distinguish \emph{walk-away} failures (the
agent ends far from a reachable goal and never returns) from \emph{false
stops} (premature, false-positive stops); a walk-away is
\emph{irreversible} when the trajectory was once closer to the goal but
never came back. Degradation shifts failures toward walk-aways: among
failed episodes, the walk-away fraction rises from $78\%$ (clean) to
$88\%$ (low-light), the irreversible fraction from $73\%$ to $79\%$
(Figure~\ref{fig:motivation}b), and the false-stop rate from $0.55$ to
$0.76$. The reliability at the turning point $t^\star$, where $d_t$
switches from decreasing to increasing, drops from $0.95$ to $0.40$. The
failures are thus not insufficient exploration; they are early, confident
commitments made when perception is untrustworthy.

\paragraph{M3--M4: Reliability is informative; a single low-$r$ frame triggers the commitment.}
The reliability estimate carries a usable signal. On clean input, bucketing
commitments by $r$ shows net geodesic progress over the next five steps
rising monotonically with $r$ ($0.133/0.154/0.245$ for low/mid/high), and
across severities $0$--$4$ the mean $r$ falls monotonically
($0.92{\to}0.87{\to}0.70{\to}0.53{\to}0.41$) while the irreversible
fraction saturates at $95\%$ by severity~$3$ (Figure~\ref{fig:motivation}c).
The commitment itself is a single-frame event: the detector's
\texttt{found} flag flips on one frame, not after temporal consensus.
Across conditions, the $r$ of that triggering frame falls from $0.95$ on
clean to $0.41$ under low-light ($85/90$ clean and $84/90$ low-light
episodes commit), and
once made the commitment is never revoked by the backbone
(Figure~\ref{fig:motivation}d). The entry point of the failure is an
irreversible decision on one unreliable frame.

\paragraph{M5: A reachability/commit-loss decomposition at the $N{=}300$ scale.}
The $M$-study above uses $n{=}90$; we now confirm the mechanism at the
$N{=}300$ main-split scale, where the baseline success rate is $0.37$ rather
than the $0.22$ of \texttt{val\_mini}. We decompose the outcome as
$\text{SR} = \text{reachability} - \text{commit/stop loss}$, where reachability
is the fraction of episodes whose trajectory passes within $1.0$\,m of the goal
(an upper bound achievable by a perfect stop policy on the \emph{actual}
trajectory; the threshold is calibrated from the $p_{90}$ stopping
distance of clean successes, $0.15$\,m). On low-light at severity~$4$,
the baseline reaches within $1.0$\,m in $61.0\%$ of episodes but succeeds in
only $37.0\%$, a commit/stop loss of $24.0$ points. A comparable
conversion gap already afflicts clean input (reach $0.69$, SR $0.44$)---the
commit/stop layer is a pre-existing weakness of the frozen agent;
degradation lowers reachability ($0.69\to0.61$) and strips the deciding
frame of the evidence a test-time fix would need to repair the conversion.
Degradation does not suppress \emph{action}: the commit rate is
$88.7\%$ under low-light versus $93.0\%$ on clean, and pure-exploration
failures stay a small minority ($7.0\%\to11.3\%$). What collapses is
commitment \emph{quality}: the mean reliability at the committing frame
falls from $0.88$ to $0.39$, and only $42\%$ of commitments lead to
success. The $n{=}90$ M-study shows the same collapse on its harder
\texttt{val\_mini} slice, and the same decomposition holds for Gaussian
noise (reachability ${\sim}0.59$): the pattern is not specific to low
light.

\paragraph{Summary.}
The observations form a chain: degradation does not blind the
agent or stop it from acting; it leads the agent to stake irreversible
commitments on single, low-reliability frames, so the loss sits in the
reach-to-success conversion rather than in perception or exploration. The
diagnosis predicts that interventions on appearance or where-to-go should
not move success, because they never touch the conversion. We translate this 
diagnosis into  interventions and evaluate them across six families; no deployable intervention improves success, and several significantly reduce it.
