\section{Conclusion}

We asked why VLM-driven navigation agents fail under visual degradation and
showed that the common remedy---restoring the input---is aimed at the wrong
layer: the agent keeps acting and keeps seeing, but commits irreversibly on
unreliable frames, leaving a 26-point reach-to-success gap that no deployable
test-time intervention closes. The sweep, its oracle upper bound, and the
cross-corruption replication together locate this loss at the commitment
interface itself rather than in any specific corruption channel or
module.
The remaining headroom is accessible only by changing what the agent was
\emph{trained} on: degradation-aware training and adaptation that keep
perception and commitment discriminative under corruption, and, where
multi-modal sensing is available, detectors that have \emph{learned} to fuse
RGB with depth under degraded input. A first pilot in this direction---a
$10^{6}$-parameter feature adapter trained on scene-disjoint
(clean,\,degraded) pairs---already matches the best deployable test-time
arm while measurably re-aligning degraded features
(Appendix~\ref{app:road1}); the effect is not yet statistically
resolvable, but it is the qualitative direction the diagnosis predicts.
We release the benchmark, paired protocol, and all negative results as an
instrument for measuring such retrained agents under the same paired,
adequately powered conditions.

\paragraph{Limitations.}
Our diagnosis is bounded by its substrate: one simulator (Habitat HM3D),
one task (ObjectNav), one open-weight VLM backbone (Qwen2-VL-7B), and
synthetic ImageNet-C-style corruptions; real sensor degradation may add artefacts (rolling shutter,
auto-exposure hunting) we do not model. The intervention sweep covers six
causal families but cannot exhaust the design space of test-time fixes, so
our negative results bound the families we probed, not every conceivable
patch. Finally, the reliability estimator is an analysis instrument
calibrated on our probe set; we do not claim it transfers unchanged to
other embodiments.
